% !TeX root = main.tex
\documentclass[11pt]{report}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{color}
\usepackage{appendix}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[T1]{fontenc}

\usepackage{tikz}
\usetikzlibrary{patterns}

\usepackage{biblatex}
\addbibresource{ref.bib}
\titleformat{\chapter}{\normalfont\huge}{\thechapter.}{32pt}{\huge}

\lstset
{
    numbers=left,
    stepnumber=1,
    showstringspaces=false,
    tabsize=1,
    basicstyle=\footnotesize\ttfamily\color{black},
    breaklines=true,
    xleftmargin=0.225in,
    frame=l,
    columns=fixed,
    basewidth=0.5em,
    literate={~}{{\fontfamily{ptm}\selectfont \textasciitilde}}1
}

\setlength{\parindent}{0pt}

\begin{document}
\include{title.tex}
\tableofcontents
\newpage

\chapter{Data Lab}
\subsection{A}
% Describe your implementation of howManyBits(x)
\begin{lstlisting}[language=C]
int howManyBits(int x) {
  // Assign integers to us later;
  int ret = 0, y = 0;

  // Get the sign of x
  // All 1s if negative, all zeroes if positive
  int sign = x >> 31;

  // If the number is negative get -x-1 
  // Aka if -12 get 11, or if 12 get 12
  // The number of digits for those 2 are exactly the same, so doing this just makes it significantly easier to work with
  x = ((sign & ~x)|(~sign & x));
 
  // Do a sort of command and conquer
  // First shift by the first half of the int size
  // Then check if there are any bits remaining
  // If there is, then set y to 16, otherwise its 0
  // Shift x by y bits, which means we wont count those bits again. If there was less than 16 bits, nothing happens.
  //  
  // Repeat this process, halfing the amount each time. 
  y = !!(x >> 16) << 4; x >>= y; ret += y;
  y = !!(x >> 8) << 3;  x >>= y; ret += y;
  y = !!(x >> 4) << 2;  x >>= y; ret += y;
  y = !!(x >> 2) << 1;  x >>= y; ret += y;
  y = !!(x >> 1);       x >>= y; ret += y; // No need to shift 1, as it would be by 0
  ret += x; // If there is more bits left, its simply 1, and is just set to x
  return 1+ret; // Return the result, adding a bit for the sign
}
\end{lstlisting}

The function $howManyBits(x)$ determines how many bits are necessary to represent the number $x$. To make the computations simpler, first get the sign of the number and negate negative numbers to one less. Eg $-12$ becomes $11$. Those 2 are represented by the same amount of bits, so it just makes things simpler. The mutations done to $-12$ are as so:
\begin{align*}
    -12 &= 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 0100\\
    \sim-12 &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 1011\\
    \text{sign}  &= 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 1111\\
    \sim\text{sign} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\\
    -12 \& \sim-1 &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\\
    \sim-12 \& \text{sign} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 1011\\
    (-12 \& \sim\text{sign}) | (~-12 \& \text{sign}) &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 1011\\
    11  &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 1011
\end{align*}
For $x=11$ the mutations look like:
\begin{align*}
    11 &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 1011\\
    \sim11 &= 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 0100\\
    \text{sign} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\\
    \sim\text{sign} &= 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 1111\ 1111\\
    11 \& \sim\text{sign} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 1011\\
    \sim11 \& \text{sign} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\\
    (11 \& \sim\text{sign}) | (~11 \& \text{sign}) &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 1011
\end{align*}
For positive numbers we change nothing in the end, but it's necessary for the algorithm to work.\\[1ex]

Next step is to do a search for the outermost bit. This is done by shifting the number by half the size of the integer, initially 16. If there are any bits left, then add 16 to the result. And shift the number by 16. Repeat with 8,4,2 and 1. Doing so will narrow us down more and more to the result. One thing to note is that we always add 1 at the end of the method for the sign. If we take 17 as an example, skipping the 16 and 8 shifts. 
\begin{align*}
    17 &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0001\ 0001\\
    \text{shift by 4} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0001\\
    \text{add 4 to y} &= 4\\
    \text{shift x by 4} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0001\\
    \text{shift by 2} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\\
    \text{shift by 1} &= 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\ 0000\\
    \text{add x to y} &= 5\\
    \text{add 1 for sign} &= 6
\end{align*}
So 17 is stored using 6 bits, 5 for the number and 1 for the sign.

\subsection{B}
\begin{lstlisting}[language=C]
int tmin(void) {
    // The absolute minimum value is 0x80000000, which can be achieved by shifting 1 left 31 bits.
    return 1 << 31;
}
\end{lstlisting}

The absolute minimum value of an int is, by two's complement, 0x80000000, aka 1 only followed by 0. This can be achieved by shifting 1 left 31 bits.


\chapter{Attack Lab}
% What happens when the c3 assembly instruction is executed? Does anything on the stack change?
\section{A}
The \textit{ret} instruction pops the return address off the stack and transfers control to it. When a \textit{call} is executed the first thing the new method does is push the old stack frame base pointer onto the stack, which is then popped at the end of the method. It's not part of the \textit{ret} instruction, but it's a part of the standard call stack. Anoteher thing not directly done by \textit{ret} but that's worth noting is that the \textit{\%eax} register, is set to the return value, this is used by the calling method to get the return value.

\section{B}
% What is a gadget farm?
A gadget farm is a section of code that contains gadgets. A gadget is a part of code that contains instructions followed by a \textit{ret} instruction. These can be used to execute small parts of code in an exploit. Such as moving a value to a register, performing small or a single operation etc. Gadgets arent't often present on purpose, they are extracted from other code. Such as setting a value to something specific,  and then returning. This might accidentally encode into something that can be used as a gadget.\\

Gadgets are, usually, not especially useful on their own. But if used together they can be a powerful tool. For example, if we have a gadget that moves a value to a register, and another that performs an operation on that register, we can use them together to perform an operation on a value.\\[1ex]

In the case of attacklab, the goal was to run a method with an argument. This meant setting the \textit{\%rdi} register to the argument. The value of the argument is on the top of the stack, so we need to find a way to pop to \textit{\%rdi}. This can't be directly done with our gadgets, but we can pop to \textit{\%rax} with this gadget:
\begin{lstlisting}[language={[x86masm]Assembler}]
0000000000402985 <setval_484>:
  402985:	f3 0f 1e fa          	endbr64 
  402989:	c7 07 58 90 90 c3    	movl   $0xc3909058,(%rdi)
  40298f:	c3                   	retq   
\end{lstlisting}
The instruction \textit{58} will pop to \textit{\%rax}. Located at $402989 + 2 = 40298B$. Then we can move the value to \textit{\%rdi} with this gadget:
\begin{lstlisting}[language={[x86masm]Assembler}]
0000000000402970 <getval_104>:
    402970:	f3 0f 1e fa          	endbr64 
    402974:	b8 48 89 c7 c3       	mov    $0xc3c78948,%eax
    402979:	c3                   	retq   
\end{lstlisting}
The instruction \textit{48 89 c7} will move the value in \textit{\%rax} to \textit{\%rdi}. Located at $402974 + 1 = 402975$.\\[1ex] 

With that information collected we can construct the payload:
\begin{lstlisting}
00 00 00 00 00 00 00 00 /* padding start */
00 00 00 00 00 00 00 00
00 00 00 00 00 00 00 00
00 00 00 00 00 00 00 00
00 00 00 00 00 00 00 00 /* padding end */
8B 29 40 00 00 00 00 00 /* gadget 1 */
88 7a 22 43 00 00 00 00 /* cookie */
75 29 40 00 00 00 00 00 /* gadget 2 */
7c 27 40 00 00 00 00 00 /* touch2 address */ 
\end{lstlisting}
It starts with the padding, then the first gadget, which will be executed first. Then the cookie, which will be popped to \textit{\%rax}. Then the second gadget, which will move the value to \textit{\%rdi}. And finally the address of the \textit{touch2} method, which will be executed after the gadget.\\[1ex]

\chapter{Malloc Lab}
\section{A}
% Explain in detail your implementation of the mm_malloc function
The function \textit{mm\_malloc}, see appendix \ref{appendix:malloc_lab_code_mm_malloc}, takes a size $x$, finds or creates a spot for it on the heap and places it, removing it from available memory. My implementation of malloclab uses an explict free list, which is essentially just a linked list between all the free blocks of memory.

An 8 word block of memory is allocated as 10 words, where 2 are used for the header and footer, figure \ref{fig:malloc_lab_a} illustrates such a free block.\\

\begin{figure}[h]
    \centering
    \begin{tikzpicture}
        \draw[thick] (0,0) -- (0,2) -- (10,2) -- (10,0) -- (0,0);
        
        \draw[thick] (1,0) -- (1,2);
        \draw[thick] (2,0) -- (2,2);
        \draw[thick] (3,0) -- (3,2);
        \draw[thick] (9,0) -- (9,2);

        \node[rotate=90] (header) at (0.5,1) {Header};
        \node[rotate=90] (next) at (1.5,1) {Next};
        \node[rotate=90] (prev) at (2.5,1) {Prev};
        \node[rotate=90] (footer) at (9.5,1) {Footer};
    \end{tikzpicture}
    \caption{Free block}
    \label{fig:malloc_lab_a}
\end{figure}

The header and footer of a block are identical, they contain the size of the block and a bit indicating allocation state. The next and prev pointers are used to link to the next and previous blocks in the free list, using their addresses.\\[1ex]

The \textit{mm\_malloc} method first checks if the heap is initialized, if not it initializes it with a call to \textit{mm\_init}. Then we align the given size to be double word aligned, and attempt to find a free block, with \textit{find\_fit} that can contain the aligned size. If not possible the heap is extended, with \textit{extend\_heap}, which creates a new free block. Then we place the block, either in the found block or the newly created block, with \textit{place}. Each of these methods are described in the following sections, as well as \textit{coalesce} which is used to merge free blocks. 

\subsection{find\_fit}
\textit{find\_fit}, appendix \ref{appendix:malloc_lab_code_find_fit}, is extremely naively implemented. It simply iterates through the free list and returns the first block that is large enough to fit the given size.\\[1ex]

This is far from the most efficient method. There is also the method of best fit which attempts to waste as little space as possible, but requires iterating the entire list. Alternatively there is next fit, which keeps track of the last block used, and searches from there. This has the advantage of not fragmenting the beginning of memory as much as first fit, but it might take a while to reuse freed memory.

\subsection{extend\_heap}
\textit{extend\_heap}, appendix \ref{appendix:malloc_lab_code_extend_heap}, calls \textit{mem\_sbrk}, an external method, to extend the heap by a given number of words. Then it creates a free block that spans the entirety of the newly allocated memory. To end with it calls \textit{coalesce} to merge the new block with any adjacent free blocks. In the case of \textit{extend\_heap} it can only coalesce with the block to it's left, but to not add complexity the existing method is used.\\[1ex]

\subsection{coalesce}
\textit{coalesce}, appendix \ref{appendix:malloc_lab_code_coalesce}, merges adjacent free blocks. It takes a pointer to a block, and checks if the previous and next blocks are free. If they are it removes them from the free list, merge with the given block and adds the combined block to the free list. In case no blocks can be merged, it adds the given block to the free list\\[1ex]

There are 4 cases that coalesce can handle, figure \ref{fig:malloc_lab_coalesce} illustrates those cases.
\begin{figure}[h]
    \begin{subfigure}[b]{\textwidth}
        \centering
        \begin{tikzpicture}
            \draw[thick, pattern=north west lines] (0,0) rectangle (4,2);
            \draw[thick] (4,0) rectangle (8,2);
            \draw[thick, pattern=north west lines] (8,0) rectangle (12,2);
        \end{tikzpicture}
        \caption{No merge}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \begin{tikzpicture}
            \draw[thick, pattern=north west lines] (0,0) rectangle (4,2);
            \draw[thick] (4,0) rectangle (8,2);
            \draw[thick] (8,0) rectangle (12,2);
        \end{tikzpicture}
        \caption{Merge with right}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \begin{tikzpicture}
            \draw[thick] (0,0) rectangle (4,2);
            \draw[thick] (4,0) rectangle (8,2);
            \draw[thick, pattern=north west lines] (8,0) rectangle (12,2);
        \end{tikzpicture}
        \caption{Merge with left}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \begin{tikzpicture}
            \draw[thick] (0,0) rectangle (4,2);
            \draw[thick] (4,0) rectangle (8,2);
            \draw[thick] (8,0) rectangle (12,2);
        \end{tikzpicture}
        \caption{Merge with both left and right}
    \end{subfigure}

    \centering
    \caption{Coalesce cases}
    \label{fig:malloc_lab_coalesce}
\end{figure}

\newpage
\subsection{place}
\textit{place}, appendix \ref{appendix:malloc_lab_code_place}, takes a pointer to a block and a size. It checks if the given block is large enough to split into two. If it is it splits it by resizing the given block to the given size, done by creating a new footer with the new size, and overwriting the size in the header. It then creates a new header after the footer, with the remaining size, and overwrites the original footer. This new block is added to the free list by calling \textit{coalesce}. If it cannot split the block \textit{place} simply sets the allocation bit to true\\[1ex]

\subsection{Adding to, and removing from the free list}
Two methods \textit{insert\_in\_empty\_list} and \textit{remove\_from\_empty\_list} provide the interface with the free list.\\[1ex]

\textit{insert\_in\_empty\_list}, appendix \ref{appendix:malloc_lab_code_insert_in_empty_list}, takes a pointer to a block and inserts it in the empty list. Due to the LIFO principle (Last in first out) the method first sets the previous pointer on the root block (first block in the list) to the given pointer, then sets the next pointer of the current block to the root block, lastly set the previous pointer of the current block to null. See \ref{fig:malloc_lab_list}\\[1ex]

\textit{remove\_from\_empty\_list}, appendix \ref{appendix:malloc_lab_code_remove_from_empty_list}, takes a pointer to a block that should be removed from the empty list. If we're the root then we set the next block's previous to null, and store it's pointer as the new root. Otherwise we set the previous block's next pointer the the next block and the previous pointer of the next block to the previous block.\\[1ex]

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{\textwidth}
        \centering
        \begin{tikzpicture}
            \draw[thick] (0,0) rectangle (4,2);
            \draw[thick] (4,0) rectangle (8,2);
            
            \draw[thick] (1,0) -- (1,2);
            \draw[thick] (2,0) -- (2,2);
            \draw[thick] (3,0) -- (3,2);

            \node[rotate=90] (header) at (0.5,1) {Header};
            \node[rotate=90] (next) at (1.5,1) {Next};
            \node[rotate=90] (prev) at (2.5,1) {Prev};
            \node[rotate=90] (footer) at (3.5,1) {Footer};

            \node (null) at (-1,3) {NULL};

            \draw[thick, ->] (1.5,2) to[out=90, in=0] (null);
            \draw[thick, ->] (2.5,2) to[out=90, in=0] (null);
        \end{tikzpicture}
        \caption{Before}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \begin{tikzpicture}
            \draw[thick] (0,0) rectangle (4,2);
            \draw[thick] (4,0) rectangle (8,2);
            
            \draw[thick] (1,0) -- (1,2);
            \draw[thick] (2,0) -- (2,2);
            \draw[thick] (3,0) -- (3,2);

            \draw[thick] (5,0) -- (5,2);
            \draw[thick] (6,0) -- (6,2);
            \draw[thick] (7,0) -- (7,2);

            \node[rotate=90] (header) at (0.5,1) {Header};
            \node[rotate=90] (next) at (1.5,1) {Next};
            \node[rotate=90] (prev) at (2.5,1) {Prev};
            \node[rotate=90] (footer) at (3.5,1) {Footer};

            \node[rotate=90] (header) at (4.5,1) {Header};
            \node[rotate=90] (next) at (5.5,1) {Next};
            \node[rotate=90] (prev) at (6.5,1) {Prev};
            \node[rotate=90] (footer) at (7.5,1) {Footer};

            \draw[thick, ->] (2.5,0) to[out=270, in=270] (5.5,0);
            \draw[thick, ->] (5.5,2) to[out=90, in=90] (1.5,2);

            \node (null) at (-1,3) {NULL};

            \draw[thick, ->] (1.5,2) to[out=90, in=0] (null);
            \draw[thick, ->] (6.5,2) to[out=90, in=0] (null);
        \end{tikzpicture}
        \caption{After}
    \end{subfigure}
    \caption{Inserting in the empty list}
    \label{fig:malloc_lab_list}
\end{figure}
\newpage
\section{B}
% What is pointer arithmethic? Describe how you used it in your version of mm.c
In C a pointer is a numeric value. Therefore we can perform arithmethic on it. Such as adding or subtracting a number to it. This is called pointer arithmetic.\\[1ex]

In my version of \textit{mm.c} I used pointer arithmetic to find the location of the header, footer and previous pointer. It was also used to find the location of the next, and previous, block in memory, calculated using the size of the block. 
\begin{lstlisting}[language=C]
// Compute the address of the header, from a pointer to the data location
#define HDRP(bp) ((char *)(bp)-WSIZE)
// Compute the address of the footer, from a pointer to the data location
#define FTRP(bp) ((char *)bp + GET_SIZE(HDRP(bp)) - DSIZE)

// Compute the location of the next block by the data location of a block
#define NEXT_BLKP(bp) ((char *)bp + GET_SIZE((char *)(bp)-WSIZE))
// Compute the location of the previous block by the data location of a block
#define PREV_BLKP(bp) ((char *)bp - GET_SIZE((char *)(bp)-DSIZE))

// Get the location of the pointer to the next block
#define NEXT_FBLKP(bp) ((char *)bp)
// Get the location of the pointer to the next block
#define PREV_FBLKP(bp) ((char *)bp+WSIZE)
\end{lstlisting}
All these usecases assume the given pointer is at the beginning of the block's data. To find the header simply remove a word, as it's placed right before the data, and you have the pointer to the header location. To find the footer add the size of the block, this is retrieved from the header, and subtract a double word, as the footer is placed right after the data\\[1ex]

Next block is found by adding the size of the block to the pointer. Previous pointer is calculated by getting the size of the previous blocks footer, stored 2 words before the data, and subtracting it from the pointer.\\[1ex]

The next free block pointer location is stored at the data location. The previous free block pointer location is stored 1 word after the data location, so simply add a word.

\chapter{Topics from class}
\section{A}
% Goldmine: https://stackoverflow.com/questions/49396346/why-are-segfaults-called-faults-and-not-aborts-if-they-are-not-recoverable
Traps are usually intentional and emitted by a software process. Let's for example take a zero division error. In this case the kernel's trap handler would send a signal to the process. Traps are reported after the instruction that caused it has executed.\\[1ex]

Faults, unlike traps, are usually unintentional and are emitted by hardware. While not intentional some faults can be recovered, fx. if a process dereferenced a pointer to paged memory. In which case the kernel's trap handler would load the paged memory into main memory to actually access the data. The kernel doesn't send a signal to the process, in this case, but just retries the instruction that caused the page fault. Faults are detected before or during the instruction, and if they are reported, it's done after the environment has been reset to a state that can retry the instruction\\[1ex]

Aborts are unintentional and unrecoverable. Fx a segfault, despite it's name, will usually kill an application. Segfaults happen when an application tries to access an invalid address. The kernel handler sends a signal to the process.\\[1ex]

One thing to note about the above is that it assumes a very simple architecture. In most modern OS's all of these are recoverable, wouldn't be very convenient if all segfaults crashed the operating system. You can actually catch aborts in C on linux systems by registering a signal handler for them, as interrupts are simply signals. 

\section{B}
Before understanding ports, we must understand the concept of sockets. A socket is an endpoint for network communication. A socket has and IP address and a port, formatted as \textit{address:port}, also known as a socket address. Sockets are used to communicate with other sockets, usually between a servers socket and a clients socket.\footnote{All information in this section is from Computer Systems: A Programmers Perspective \cite{computer-systems}, reworded by me.}\\[1ex]

Ephemeral ports are short lived ports, usually used for the client to server communication. When a client tries to make a request to the server the kernel will assign a port to the socket address used. Once the client disconnects from the server, this port is closed.\\[1ex]

A well known port is a long lived ports, usually used for the server to client communication. Well known ports are usually one of the permanently assigned ports. Fx port 80 is reserved for HTTP traffic, it can be used for other purposes, but will consfuse HTTP clients. 

\section{C}
A memory leak is allocated memory that is no longer used. It can occur when one forgets to clean up the memory they have allocated or when allocating memory and overwriting the pointer to it, without first releasing the memory (orphaning). It can be avoided by using garbage collection, or by remembering to always have a \textit{free} call for every \textit{malloc} call.\\[1ex]

Garbage collection is a technique for freeing memory that is no longer used by the program. A garbage collector holds a reachability tree, which contains root nodes and heap nodes. If there is no path to a heap node, from some root node, it is considered garbage and can be freed. In C we use a so called conservative garbage collector, which assumes that all blocks are still in use, unless proven otherwise. This does cause some blocks to be incorrectly identified as in use, but ensures no blocks are incorrectly identified as garbage.\\[1ex]

\section{D}
Parallel processing is the act of exeucting multiple instructions at the same time, instead of one at a time. Unfortunately, there are no guarantees that the instructions will be executed in the order they were written. Therefore when multiple threads are modifying the same data, the order of modifications are not guaranteed. This problem is called race conditions.\\[1ex]

Due to the unpredictable nature of race conditions they are hard to debug. They happen semi randomly and are therefore hard to replicate. It all depends on how the computations are scheduled by the CPU.\\[1ex]

To avoid race conditions synchronization is needed.
One synchronization primitivie is mutexes which, when locked, are only locked on that thread. These can be used to ensure only 1 thread have access to the data. The problem with this approach is that it can make the code synchronous if used inproperly, or become a bottleneck. On page 1050 of Computer Systems \cite{computer-systems} an experiment is conducted. This experiment uses a varying amount of threads to sum multiple numbers, using mutexes to lock the result variable. In this experiment threading is orders of magnitude slower than single threaded due to the massive overhead of mutex locking.\\[1ex]

Another approach is the \textit{\_thread} keyword, which creates an instance of the variable pr. thread. That's expensive in terms of memory, and pretty useless if the data needs to be shared, such as multithreaded computations on the same data.\\[1ex]

Said best by Randal E. Bryant and David R. O'Hallaron \textit{"Synchronization overhead is expensive and should be avoided if possible."} \cite{computer-systems}. If possible synchronization should be completely avoided, or used sparingly. Do as many operations as possible outside the mutex, such that the time spent in the locked mutex is as small as possible.\\[1ex]

\printbibliography

\chapter{Appendix}
\appendix
\chapter{Malloc Lab code}
\section{mm\_malloc}\label{appendix:malloc_lab_code_mm_malloc}
\begin{lstlisting}[language=C]
void *mm_malloc(size_t size) {
    size_t asize;      // Adjusted block size 
    size_t extendsize; // Amount to extend heap if no fit 
    void *bp;
    
    if (heap_listp == 0) {
        mm_init();
    }
    
    // Ignore spurious requests 
    if (size == 0)
        return NULL;
    
    asize = get_alligned(size);
    if ((bp = find_fit(asize)) == NULL) {
    
        extendsize = MAX(asize, CHUNKSIZE);
        if ((bp = extend_heap(extendsize / WSIZE)) == NULL)
        return NULL;
    }
    
    // No fit found. Get more memory and place the block 
    place(bp, asize);
    
    return bp;
}
\end{lstlisting}

\section{find\_fit}\label{appendix:malloc_lab_code_find_fit}
\begin{lstlisting}[language=C]
static void *find_fit(size_t asize)
{
  // First-fit search 
  void *bp = first_freep;

  Brendan Fairclough

  while (bp != NULL) {
    if (GET_SIZE(HDRP(bp)) >= asize)
      return bp;
    bp = NEXT_FBLK(bp);
  }

  return NULL;
}
\end{lstlisting}

\section{extend\_heap}\label{appendix:malloc_lab_code_extend_heap}
\begin{lstlisting}[language=C]
    static void *extend_heap(size_t words) {
  char *bp;
  size_t size;

  // Allocate an even number of words to maintain alignment 
  size = (words % 2) ? (words + 1) * WSIZE : words * WSIZE;
  if ((long)(bp = mem_sbrk(size)) == -1)
    return NULL;

  // Initialize free block header/footer and the epilogue header 
  // Overwrites old epilogue header, notice HDRP
  PUT(HDRP(bp), PACK(size, 0));         // Free block header 
  PUT(FTRP(bp), PACK(size, 0));         // Free block footer 
  PUT(HDRP(NEXT_BLKP(bp)), PACK(0, 1)); // New epilogue header 

  // Coalesce if the previous block was free 
  return coalesce(bp);
}
\end{lstlisting}

\section{coalesce}\label{appendix:malloc_lab_code_coalesce}
\begin{lstlisting}[language=C]
static void *coalesce(void *bp) {
  // Is the previous block allocated
  size_t prev_alloc = GET_ALLOC(FTRP(PREV_BLKP(bp)));
  // Is the next block allocated?
  size_t next_alloc = GET_ALLOC(HDRP(NEXT_BLKP(bp)));
  // Get the size of the current block
  size_t size = GET_SIZE(HDRP(bp));

  // Sandwiched between 2 allocated blocks
  if (prev_alloc && next_alloc) {
    insert_in_empty_list(bp);
    return bp;
  }
  
  // Previous is allocated, but the next is free
  else if (prev_alloc && !next_alloc) {
    // Get the combined size of the current and next block
    size += GET_SIZE(HDRP(NEXT_BLKP(bp)));
    // Remove from empty list
    remove_from_empty_list(NEXT_BLKP(bp));
    // Overwrite the current block size
    PUT(HDRP(bp), PACK(size, 0));
    PUT(FTRP(bp), PACK(size, 0));
  }

  // Previous is free, but the next is allocated
  else if (!prev_alloc && next_alloc) {
    // Get the combined size of the current and next block
    size += GET_SIZE(HDRP(PREV_BLKP(bp)));
    // Remove from empty list
    remove_from_empty_list(PREV_BLKP(bp));
    // Overwrite the header of the previous block
    PUT(FTRP(bp), PACK(size, 0));
    PUT(HDRP(PREV_BLKP(bp)), PACK(size, 0));
    // Return the bp of previous blocks original position
    bp = PREV_BLKP(bp);
  } 

  // Both are free
  else {
    // Get the full size
    size += GET_SIZE(HDRP(PREV_BLKP(bp))) + GET_SIZE(FTRP(NEXT_BLKP(bp)));
    // Remove from empty list
    remove_from_empty_list(NEXT_BLKP(bp));
    remove_from_empty_list(PREV_BLKP(bp));
    // Overwrite the header of the previous block
    PUT(HDRP(PREV_BLKP(bp)), PACK(size, 0));
    // Overwrite the footer of the next block
    PUT(FTRP(NEXT_BLKP(bp)), PACK(size, 0));
    // Return the bp of previous blocks original position
    bp = PREV_BLKP(bp);
  }
  insert_in_empty_list(bp);

  // No change so return the current block
  return bp;
}
\end{lstlisting}

\section{place}\label{appendix:malloc_lab_code_place}
\begin{lstlisting}[language=C]
static void place(void *bp, size_t asize)
{
    // Get the size of the block
    size_t csize = GET_SIZE(HDRP(bp));

    remove_from_empty_list(bp);
    // Split if there is space for another block, and its headers after our data
    if ((csize - asize) >= (2 * DSIZE)) {
    // Create the block for our data and allocate it
    PUT(HDRP(bp), PACK(asize, 1));
    PUT(FTRP(bp), PACK(asize, 1));
    // Create pointer for the block after
    bp = NEXT_BLKP(bp);
    // Create new free block
    PUT(HDRP(bp), PACK(csize - asize, 0));
    PUT(FTRP(bp), PACK(csize - asize, 0));

    coalesce(bp);
    } else {
    // Set the block as allocated
    PUT(HDRP(bp), PACK(csize, 1));
    PUT(FTRP(bp), PACK(csize, 1));
    }
}
\end{lstlisting}

\section{insert\_in\_empty\_list}\label{appendix:malloc_lab_code_insert_in_empty_list}
\begin{lstlisting}[language=C]
static void insert_in_empty_list(void *bp) {
  set_prev_fblkp(first_freep, bp);
  set_next_fblkp(bp, first_freep);
  set_prev_fblkp(bp, NULL);

  first_freep = bp;
}
\end{lstlisting}

\section{remove\_from\_empty\_list}\label{appendix:malloc_lab_code_remove_from_empty_list}
\begin{lstlisting}[language=C]
static void remove_from_empty_list(void *bp) {
    void *prevp = PREV_FBLK(bp);
    void *nextp = NEXT_FBLK(bp);
  
    if (prevp == NULL) {
      set_prev_fblkp(nextp, NULL);
      first_freep = nextp;
    } else {
      set_next_fblkp(prevp, nextp);
  
      if (nextp != NULL)
        set_prev_fblkp(nextp, prevp);
    }
  
    set_next_fblkp(bp, 0);
    set_prev_fblkp(bp, 0);
}
\end{lstlisting}

\end{document}